### 2021-01-22 docs/manuscript.tex
resolve some remarks in the manuscript

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[fleqn,11pt]{SelfArx} % Document font size and equations flushed left
\usepackage[justification=justified]{caption}
\usepackage{parskip}
\usepackage{microtype}
\usepackage{svg}
\usepackage{float}
\usepackage{placeins}
\usepackage[english]{babel} % Specify a different language here - english by default
\usepackage{listings}
\usepackage{fontspec}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{makecell}
\usepackage{soul}
% \usepackage{lipsum} % Required to insert dummy text. To be removed otherwise
\setmainfont{TeX Gyre Termes}
%----------------------------------------------------------------------------------------
%	COLUMNS
%----------------------------------------------------------------------------------------

\setlength{\columnsep}{0.55cm} % Distance between the two columns of text
\setlength{\fboxrule}{1pt} % Width of the border around the abstract

%----------------------------------------------------------------------------------------
%	COLORS
%----------------------------------------------------------------------------------------

\definecolor{color1}{RGB}{124,0,0} % Color of the article title and sections
\definecolor{color2}{RGB}{150,150,150} % Color of the boxes behind the abstract and headings

%----------------------------------------------------------------------------------------
%	HYPERLINKS
%----------------------------------------------------------------------------------------

\usepackage{hyperref} % Required for hyperlinks
\hypersetup{hidelinks,colorlinks,breaklinks=true,urlcolor=color2,citecolor=color1,
  linkcolor=color1,bookmarksopen=false,pdftitle={Title},pdfauthor={Author}}

%----------------------------------------------------------------------------------------
%	ARTICLE INFORMATION
%----------------------------------------------------------------------------------------

\PaperTitle{
    HAMdetector: A Bayesian regression model that integrates information to detect HLA-associated mutations
} % Article title

\Authors{Daniel Habermann\textsuperscript{1}, ..., Daniel Hoffmann\textsuperscript{1}} % Authors
\affiliation{\textsuperscript{1}\textit{Bioinformatics \& Computational Biophysics, Faculty of Biology, University of Duisburg-Essen, 45117 Essen, Germany}} % Author affiliation
% \affiliation{\textsuperscript{2}\textit{Department of Chemistry, University of Examples, London, United Kingdom}} % Author affiliation
% \affiliation{*\textbf{Corresponding author}: john@smith.com} % Corresponding author

\Keywords{human leukocyte antigen system, HLA, multiple sequence alignment, escape mutations, 
  viral escape, Bayesian inference, sparsity, horseshoe, epitope prediction} % Keywords - if you don't want any simply remove all the text between the curly brackets
\newcommand{\keywordname}{Keywords} % Defines the keywords heading name

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

\Abstract{
  \\
  \textbf{Motivation} \\
  The human leukocyte antigen system (HLA) presents viral peptide fragments in MHC I protein-peptide complexes on cell surfaces and in this way alerts CD8\textsuperscript{+} cytotoxic T-Lymphocytes (CTLs), a key component of adaptive immunity. This pathway exerts a strong selection pressure on viruses, favoring viral mutants that escape recognition by the HLA/CTL system, e.g.\ by point mutations that decrease binding of viral peptides to MHC I. Naturally, such immune escape mutations often emerge in highly variable viruses, e.g.\ HIV or HBV, as HLA-associated mutations (HAMs), specific to the host HLA alleles and its MHC I proteins. The reliable identification of HAMs is not only important to understand viral genomes and their evolution, but it also impacts the development of broadly effective anti-viral treatments and vaccines against variable viruses.\\

  By their very nature HAMs are amenable to detection by statistical methods in paired sequence / HLA data. However, HLA alleles are very polymorphic in the human host population which makes the available data relatively sparse and noisy. Under these circumstances, one way to optimize HAM detection is to integrate all relevant information in a coherent model. Bayesian inference offers a principled approaches to achieve this.\\

  \textbf{Results} \\
  Here, we present a Bayesian regression model that is easily extensible to include information from different sources, and that makes use of recent advances in Bayesian inference, e.g.\ by using a sparsifying prior. We show that this integrative approach improves performance in HAM detection considerably over state-of-the-art methods. \\

  \textbf{Availability and Implementation} \\
  The source code of this software is available at \url{http://github.com} under a permissive MIT license. \\

  \textbf{Supplementary information} \\
  \href{https://google.com}{Supplementary data} are available at \textit{Bioinformatics} online. \\
}

%----------------------------------------------------------------------------------------
\begin{document}

\flushbottom % Makes all text pages the same height

\maketitle % Print the title and abstract box

{
  \hypersetup{linkcolor=black}
  \tableofcontents
}

\section{Introduction}

\subsection{The HLA system}

One way of how the human immune system recognizes intracellular viral infections is through the human leukocyte antigen system~\cite{Germain1994}: In cells with active protein biosynthesis, proteins are continuously synthesized and also degraded by proteasomal degradation, which cleaves proteins into peptides~\cite{Goldberg2002}.

A small subset of these peptides is presented on the cell surface by receptors called MHC class I encoded by HLA genes, e.g.\ in humanc HLA-A, -B, and -C. HLA genes are very polymorphic between human individuals, with more than 20,000 different described HLA alleles~\cite{Robinson2014}. The resulting gene products differ in their binding properties, which means that cells from different individuals present a highly diverse set of peptides on the cell surface.

Cytotoxic T-Lymphocytes (CTLs) are selected during maturation to only weakly bind to MHC I protein-peptide complexes if the peptide originates from proteins of the self-proteome, but might be able to strongly bind to MHC I protein-peptide complexes with peptides generated from viral proteins~\cite{Murata2007}. Upon activation, T cells induce cytolytic activity and recruit other immune cells~\cite{Harty2000}.

\subsection{HLA escape}
In this way, the HLA system exerts strong selection pressure towards virus variants that escape T cell recognition~\cite{Borrow1997}. Such variants could, for example, carry mutations that result in reduced binding of immunogenic peptides to MHC I or to T cell receptors, or that alter the processing and binding so that peptides are no longer presented at all~\cite{Yewdell2002}.

The evolutionary events are complex and occur not only on the level of individuals, where a virus adapts to specific features of the host, but also on the population level, for example because HLA alleles differ in their frequency across geographic regions~\cite{Kawashima2009}. Upon transmission to a new host with different HLA alleles, HLA escape mutations can revert to their wild type, as they are often associated with a reduction in viral replication capacity~\cite{Matthews2008}. If the latter is not the case, such mutations may not revert but be retained in the viral genome~\cite{Kawashima2009}.

Whether and how quickly a given escape mutation is selected in a host depends on several factors, such as the viral genomic background and diversity, the magnitude of the reduction in viral replicative capacity, the availability of compensatory mutations that recover fitness, and the strength of selection pressure~\cite{Kloverpris2016}.

Immune escape is a driver of viral evolution in populations and patients, and therefore of eminent importance for the understanding of many aspects of highly variable viruses such as HIV or HBV~\cite{Alizon2011}\cite{Allen2005}\cite{Rousseau2008}\cite{Lumley2018}. This includes the development of treatments and vaccines that rely on effective immune responses. This makes detection of immune escape mutations critical.

\subsection{Identifying HLA escape mutations}

There are several experimental methods available to study HLA escape~\cite{Timm2004}. However, these methods are relatively slow and costly, especially for screening purposes. Therefore, a promising approach is to combine viral genome sequencing, host HLA determination, computational identification by statistical association analysis, and targeted experimental validation~\cite{Carlson2012}.

As the selection pressure exerted by cytotoxic T cells depends on successful recognition of viral peptides on the cell surface, escape mutations are often HLA-allele specific and can therefore be detected as HLA-allele dependent footprints in sequence alignments of viral proteins~\cite{Moore2002}: At certain alignment positions, a replacement might be more frequently observed in sequences from hosts with a specific HLA allele than in  sequences from hosts without that HLA allele. By quantifying this difference for all replacement and HLA allele pairs, it is possible to identify replacements that are enriched in sequences coming from hosts with certain HLA alleles, and thus are likely to be HLA escape mutations.

One way of quantifying this enrichment is Fisher's exact test~\cite{Fisher1922}: For a given replacement \(R_{i}\) at alignment position \(i\) and HLA allele \(H\), a 2-by-2 contingency table is constructed containing the absolute counts of the number of sequences in the four possible categories  (\(R_{i}\), \(H\)), (\(R_{i}\), \(\neg H\)), (\(\neg R_{i}\), \(H\)) and (\(\neg R_{i}\), \(H\)), where \(\neg R_{i}\) denotes any replacement except \(R_{i}\), and \(\neg H\) denotes any HLA allele except \(H\).

Fisher's exact test is a conventional null hypothesis significance test (NHST) that generates p-values. In this case, the null hypothesis is that HLA allele \(H\) and  replacement \(R_{i}\) are independent, and the p-value is the probability of observing a deviation from independence that is at least as extreme as in the data at hand under the assumption that the null hypothesis is true.

Fisher's exact test has the advantage of being easy to apply~\cite{Budeus2016},  but also has several disadvantages~\cite{Carlson2008}. The most striking one is that viral sequences share a common phylogenetic history, and,  therefore, treating sequences as independent and identically distributed samples may under- or overestimate effect sizes. In the context of hypothesis testing, this leads to increased false positive and false negative rates.

Another issue of applying Fisher's exact test is that HLA class I loci are located in proximity on chromosome 6 and are therefore in linkage disequilibrium, which means that HLA alleles are not inherited completely independent of each other, i.e. inheritance of one HLA allele correlates with inheritance of another HLA allele. When using a statistical method that tests each HLA allele individually without considering the whole set of alleles present, spurious associations might occur: If HLA allele \(H_{1}\) is associated with an amino acid replacement \(R\), but \(H_{1}\) is in linkage disequilibrium with another allele \(H_{2}\), this also means that we observe an association between replacement \(R\) and \(H_{2}\), even in absence of any underlying escape mechanism.

Correlations can not only occur between HLA alleles, but also between replacements. This kind of codon covariation occurs for example in compensatory mutations that attenuate the negative impact of immune escape mutations. For instance, a compensatory mutation might lead to a conformational change in such a way that the mutated protein resembles the original wild type more closely.

Carlson et al. \nolinebreak\cite{Carlson2008} developed a method called  Phylogenetic Dependency Network that accounts for phylogenetic bias,  HLA linkage disequilibrium, and codon covariation and is based on null hypothesis significance testing.

\subsection{Issues with using p-values as a screening tool}

In addition to the aforementioned biological reasons why Fisher's exact test is  not suited for the analysis of sequence data, there are also more fundamental statistical issues that universally occur when using p-values as a screening tool~\cite{Amrhein2017}:

In the presence of small effect sizes and high variance between measurements, as  it is typically the case when working with biological data, statistically significant results can often be misleading and are likely to be in the wrong direction (a so-called type S error) or greatly overestimate an effect (a so-called type M error)~\cite{Gelman2014}. This problem has recently been widely appreciated in the literature in the context of the current “replication crisis“, which describes the circumstance that scientific claims with seemingly strong statistical evidence fail to replicate~\cite{Ioannidis2005}.

These problems are exacerbated if the p-values are used for screening purposes (multiple testing problem). The probability  of  obtaining a statistically significant result increases with each additional test, even in absence of any real effect. When using p-values as a filter, it is therefore likely to obtain significant effects that are in fact not real. To circumvent this problem, a common strategy is to control the false discovery rate, which is the expected proportion of false positives~\cite{Benjamini1995}. These adjustment procedures have the problem that, when performing many of such comparisons, none but the very largest effects remain.

Instead of performing many hypothesis tests and trying to adjust for them, we instead prefer to fit a single, multilevel model that contains all comparisons of interest. When using multilevel models, the problem of multiple comparisons can disappear entirely and  also yield more valid estimates~\cite{Gelman2012}.

\section{Material and methods}

When possible, we choose to fit Bayesian models. By using prior information, adding problem-specific structure and partial pooling, the accuracy of estimates can often be noticeably improved~\cite{Gelman2010}. Prior information does not necessarily mean to use external data, even a rough idea about the expected magnitude of estimates is often surprisingly effective.

Additionally, Bayesian statistics provides an accessible way to test models: By comparing data generated under the model's assumptions to the actually observed data, it is possible to identify important aspects of the dataset that the models fails to capture and subsequently improve the model until it is consistent with the observed data~\cite{Gabry2019}.

In the context of identifying HLA-associated mutations we propose to improve existing methods by the following additions, which can be broadly divided into  additional information and model structure:
\subsection*{Additional information}

  \textit{Binding affinity prediction}

    HLA-associated mutations are expected to lie more frequently in regions of known epitopes. There are vast experimental binding affinity data available of different peptides and MHC I molecule pairs, and there are well-established computational methods that use these data to extrapolate HLA binding for untested peptides. We show that, by including the outcome of these computational tools as input for a probabilistic model, the prediction of HLA-associated mutations can be improved.

  \textit{Antigen processing prediction}

    Similarly, there are also HLA-allele independent effects like antigen processing that influence presentation on MHC I. Second generational tools like MHCFlurry 2.0~\cite{ODonnell2020} use both binding affinity and antigen processing data to improve epitope prediction. For our tool HAMdetector, we use the output of MHCFlurry to benefit from this binding affinity and antigen processing data in order to predict HLA-associated mutations.

\subsection*{Model structure}

  \textit{Sparsity-inducing priors}

  Recent advantages in Bayesian inference include so-called sparsity-promoting priors. Sparsity-promoting priors convey the a priori expectation that most coefficients in a regression model are close to 0. This assumption also applies to HLA-associated mutations, because the number of epitopes that are restricted by a given HLA allele is typically small compared to the number of all possible epitopes. If no epitope spanning a given alignment position is presented on the MHC I molecule, any association between a replacement and the respective HLA allele  is likely to be due to random variation alone (or HLA linkage disequilibrium).

  It has been shown that using a sparsity-promoting prior when non-zero coefficients are sparse can drastically improve predictive performance, because the model is better able to differentiate between signal and noise.

  We show that a simple logistic regression model with a sparsity-promoting prior on the regression coefficients for the HLA alleles alone already performs roughly on-par  with the much more elaborate Phylogenetic Dependency Network approach from Carlson et al. when using the fraction of identified escapes that lie in known HLA epitopes as a benchmark.

  % \textit{Partial pooling of 4-digit HLA alleles}

  % It has recently been appreciated that binding specificities can vary drastically across  HLA alleles from the same allele group, e.g. between HLA-B*51:01 and HLA-B*51:03. For predicting HLA associated mutations, there has consequently been a shift to use 4-digit resolution data whenever available. This is not without downsides however, because overall, binding specificities are more similar for alleles in the same group, and, therefore, treating them as completely separate might unnecessarily fragment the available data.

  % In Bayesian statistics, it is not required to chose one extreme or the other: Instead of either treating all HLA alleles from the same allele group as identical or as completely separate, partial pooling of the estimates allows something in between: With partial pooling, estimates do share some information across each other, but they are also allowed to vary if necessary.

  % In the context of our model this means that observing a relevant association between a replacement and an HLA allele also makes the model more inclined to estimate relevant associations between that replacement and all other alleles from that allele group. The degree of this partial pooling can be estimated from the data, so if we do observe strong similarity of HLA alleles in a given allele group the estimates are more influenced by each other than if they do not.

\section{Implementation}

\subsection*{Model backbone}

We choose a logistic regression model as a backbone for two reasons: First, because it is easily extensible and secondly, because the coefficients are interpretable as expected increases on the log-odds scale, identical to the familiar interpretation of standard logistic regression models.
We also choose to fit the model in a Bayesian fashion using Stan, mostly because it allows us to make use of the hierarchical structure in the data and the possibility to use posterior predictive checks as a way of model checking.
Consider the following notation:

\begin{align}
  y_{ik} & \sim \text{bernoulli}(\theta_{ik}) \\
  \theta_{ik} & = \text{inv\_logit}(\beta_{0_{k}} + \sum_{j=1}^{D} X_{ij}\beta_{\text{HLA}_{jk}}),
\end{align}

where \(y_{ik}\) is the binary encoded observation at sequence \(i\) for replacement \(k\), \(\theta_{ik}\) is the estimated probability that we observe replacement \(k\) in sequence \(i\), \(X_{ij}\) is 1 if sequence \(i\) is annotated with HLA allele \(j\) and 0 otherwise, \(\beta_{\text{HLA}_{jk}}\) is the HLA regression coefficients for HLA allele \(j\) at replacement \(k\), \(D\) is the number of HLA alleles in the dataset, and \(\text{inv\_logit}\) is the logistic link function. Note that, if we observe N different amino acid states at a certain alignment position, each of those are considered in turn to be the “replacement”.

The regression coefficients \(\beta_{\text{HLA}_{jk}}\) are the main focus of interest, as they quantify the association between the occurrence of replacement \(k\) and each of the observed HLA alleles. They are in units of log-odds, so a regression coefficient of somewhere around 2 means that when comparing sequences with and without that particular HLA allele, the log-odds of observing replacement \(k\) in presence of that HLA allele is increased by 2.

Reasoning about coefficients on the log-odds scale can sometimes be unintuitive. A useful approximation to interpret logistic regression coefficients on the probability scale is the so-called divide-by-4 rule, which means that a regression coefficient of 2 corresponds to an expected increase on the probability scale of about 2/4 = 50\%.

\subsection*{Including phylogeny}

One problem that occurs when analyzing sequence data is that species share a common phylogenetic history. Standard statistical methods usually assume samples to be independent and identically distributed, which may lead to wrong conclusions when this assumption is strongly violated.
In the context of identifying HLA associations, Bhattacharya et al.~\cite{Bhattacharya2007} demonstrated the importance of correcting for the phylogenetic structure.
There are many approaches described in the literature for phylogenetic regression for binary dependent variables~\cite{Ives2014}, where the most popular approach is to estimate an additional multivariate normally distributed intercept, where the covariance matrix is based on the branch lengths of a given phylogenetic tree~\cite{Ives2009}.

This approach showed to be too computationally expensive to include in our model, so we chose a strategy that is similar to the one in Carlson et al. \cite{Carlson2008} for the Phylogenetic Dependency Network:

Consider a phylogenetic tree \(\Psi\) obtained from standard maximum likelihood methods for a given multiple sequence alignment. We are interested in estimating \(P(y_{ik}=1|\Psi)\), that is, the probability of observing the replacement \(k\) in sequence \(i\) based on the underlying phylogenetic model.
A quantity that can be readily computed using phylogenetic software like RAxML-NG~\cite{Kozlov2019} is \(P(\Psi|y_{ik}=1)\). For this, we keep the tree topology fixed, annotate the tree with the binary observations \(y\) at its leaves and optimize the branch lengths. \(P(\Psi|y_{ik}=1)\) is then the likelihood of the annotated phylogenetic tree. We can then use Bayes' rule to invert that conditional probability by additionally computing \(P(\Psi|y_{ik}=0)\), which is done by altering the annotation for sequence \(i\) and keeping all other observations constant.

In order to check if the inferred probabilities \(P(y_{ik}=1|\Psi)\) faithfully reflect the observed data we use calibration plots (see fig.~\ref{fig:phylogeny-calibration}). All observations are binned according to their inferred probability \(P(y_{ik}=1|\Psi)\). If these probabilities are calibrated correctly, we expect that in a group of observations with an inferred probability \(P(y_{ik}=1|\Psi)\) of around x\% we really do observe \(y_{ik}=1\) around x\% percent of the time.

\begin{figure}[!ht]
  \includegraphics[width=1\linewidth]{plots/phylogeny_calibration.pdf}
  \caption{Calibration plot for the HBV dataset. For a description of all datasets used in this study see section~\ref{sec:results}. Calibration plots for the other datasets are shown in the supplementary.
  All observations are first sorted by increasing estimated probability \(P(y_{ik}=1|\Psi)\) and then grouped into \(n\) bins.
  For each bin, the fraction of observations with \(y_{ik}=1\) (observed event percentage) is compared to the midpoint of each bin (the value in the center between the lowest and highest probability). The error bars show the cutpoints for each bin. If the probabilities are calibrated correctly, each dot is supposed to scatter closely around the diagonal line.}
  \label{fig:phylogeny-calibration}
\end{figure}

The estimated probabilities are then included in the model as additional intercepts:

\begin{equation}
\begin{aligned}
  y_{ik} & \sim \text{bernoulli}(\theta_{ik}) \\
  \theta_{ik} & = \text{inv\_logit}(\beta_{0_{k}} + \gamma~\text{logit}(P(y_{ik}=1|\Psi)) \\ \hspace{-6cm} &~~~~ + \sum_{k=1}^{D}X_{ik}\beta_{\text{HLA}_{jk}})
\end{aligned}
\end{equation}

The logit transform is used because is cancels out with the logistic link function. This means that the phylogeny information acts as a baseline in absence of any HLA effects. Note that we also include an additional parameter \(\gamma\), that is constrained to be positive. This helps because the inferred probabilities \(P(y_{ik}=1|\Psi)\) might not necessarily reflect the true underlying phylogenetic signal, for example because the phylogenetic tree does not match the observed data well enough.

A straight-forward extension of our model would be to also account for these sources of uncertainty, for example by using a Bayesian method to estimate a posterior distribution over possible tree topologies. The uncertainty over the tree topology and the underlying parameters of the phylogenetic model would then propagate into uncertainty of the estimated probabilities \(P(y_{ik}=1|\Psi)\). However, in order to not increase the runtime of the model further we  use standard maximum likelihood estimates and then include these in a way that additionally allows for measurement error.

\subsection*{Including HLA linkage disequilibrium and codon covariation}

The other two important confounding effects that are recognized in Carlson et al.~\cite{Carlson2008} are HLA linkage disequilibrium and codon covariation.
HLA linkage disequilibrium describes the observation that HLA alleles that lie in proximity to each other on the chromosome are also more likely to be inherited together. When using a statistical method that evaluates each HLA allele individually, this may lead to spurious associations: If some HLA allele \(X\) is associated with replacement \(k\), and HLA allele \(Y\) often co-occurs with HLA allele \(X\), this means that we also identify an association between HLA allele \(Y\) and replacement \(k\).

This association is correct in a statistical sense, but if the purpose of the model is to identify sites under HLA-mediated selection pressure, these associations due to linkage disequilibrium are most likely unwanted.

Note that these spurious associations only occur when testing each HLA allele individually. In a regression model where all HLA alleles of a sample are evaluated at once, the marginal posterior distributions for correlated predictors become broader to accurately reflect that the model is unable to discern which of these predictors explain the observed association with a replacement.

We do not currently account for codon covariation in our model, as we are interested in identifying all HLA associated mutations (which include for example compensatory mutations). The general structure of the model, in particular the sparsity assumption introduced in section~\ref{sec:sparsity}, would allow adding codon covariation by adding all other codons as additional predictors. This would generally not be possible in a frequentist setting because the predictor matrix would become singular and the inverse does not exist.

\subsection*{Including sparsity assumptions} \label{sec:sparsity}

In the context of identifying HLA associations we expect most of the HLA coefficients to be very close to 0, as a given alignment position is typically only restricted by none or few HLA alleles.
This is important prior knowledge that should be incorporated in the model.

For Bayesian models, there are a variety of different sparsity-promoting priors with slightly different properties. They share the common structure of placing most probability mass very close to 0, with large to tails to accommodate the non-zero coefficients.
For our model, we use a so-called horse prior~\cite{Carvalho2010}, which is defined as a scale-mixture of gaussians:

\begin{equation}
  \begin{aligned}
    \beta_{jk} &\sim \text{Normal}(0, \tau^{2}\lambda^{2}_{jk}) \\
    \lambda_{jk} &\sim \text{Cauchy}^{+}(0, 1) \\
    \tau &\sim \text{Cauchy}^{+}(0, \tau_{0})
  \end{aligned}
\end{equation},
where \(\beta_{jk}\) are the regression coefficients, \(\tau\) is the so-called global shrinkage parameter, \(\lambda_{jk}\) are the so-called local shrinkage parameters and \(\text{Cauchy}^{+}\) is the positively constrained Cauchy distribution.

Intuitively speaking, the global shrinkage parameter \(\tau\) is typically very small and shrinks most of the regression coefficients to 0, whereas the local shrinkage parameters can occasionally be very large to allow some coefficients to escape that shrinkage.

A useful addition to the horseshoe prior is given in Piironen et al.~\cite{Piironen2017}, which describes a way to choose the parameter governing the overall degree of sparsity \(\tau_{0}\) based on the expected number of non-zero coefficients.

For our HLA model, we utilize an extension to the horseshoe prior called the regularized horseshoe prior~\cite{Piironen2017}, which also provides some regularization for the non-zero coefficients. This is particularly useful for logistic regression models, as some shrinkage helps to deal with issues of separability and collinearity that commonly occur when fitting logistic regression models.

Figure~\ref{fig:horseshoe-comparison} shows a comparison of marginal posteriors for regression coefficients with and without horseshoe prior.

Supplying information about the expected degree of sparsity with the horseshoe prior is a considerable improvement over a standard logistic regression model  (section~\ref{sec:results}).

\begin{figure}
  \label{fig:horseshoe-comparison}
\end{figure}

\subsection*{Including epitope prediction software}

There is vast experimental data on epitope binding motifs available, e.g. through the use of mass spectrometry~\cite{Hunt1992}. Epitope prediction software like MHCFlurry~\cite{ODonnell2020} use these data to extrapolate MHC I binding for new, untested peptides.
As HLA-associated mutations are much more likely to lie in epitopes (e.g. because of selection pressure to escape MHC I binding), using epitope prediction software can provide valuable external data for identifying HLA-associated mutations.

For our model, we use MHCFlurry 2.0 for binding prediction, as it does not only predict HLA binding but also antigen processing. For this, we create an input matrix of dimensions \(R\times D\), where R is the number of evaluated replacements and D is the number of observed HLA alleles in the dataset. The elements of this matrix are binary encoded and contain a 1 if that position is either expected to be inside a predicted epitope or related to antigen processing, and 0 otherwise. Given an amino acid sequence, MHCFlurry 2.0 provides a list of possible epitopes (9-13 mers) and HLA allele pairs and calculates a rank based on comparisons with random epitope and HLA pairs. For the binarization we use a rank threshold of 0.2\% (as suggested by MHCFlurry).

We typically try to avoid thresholds whenever possible, but in this case we deemed it to be preferable to including the predicted binding affinities, as we expected the binary values to be more reliable than the predicted numeric values.
Note that the output of these epitope prediction tools may be imprecise or even wrong, but by including it as an input for a probabilistic model, the model is able to learn how much to trust this data if parameterized in a way to allow for measurement error.

We use epitope prediction as information about the expected degree of sparsity, i.e. when knowing a certain position is restricted by a certain HLA allele, we expect that this HLA allele is more likely to be associated with the replacement than the other HLA alleles.
This can be reflected by increasing the scale of the local shrinkage parameters \(\lambda_{jk}\):

\begin{equation}
  \begin{aligned}
    \lambda_{jk} &\sim \text{Cauchy}^{+}(0, \text{exp}(Z_{jk}\beta_{\text{epi}})) \\
    \beta_{\text{epi}} &\sim \text{Normal}^{+}(1, 2),
  \end{aligned}
\end{equation}

where \(Z_{jk}\) is 1 if HLA allele \(j\) is predicted to restrict the alignment position corresponding to replacement \(k\) or is associated antigen processing and 0 otherwise.
The parameter \(\beta_{\text{epi}}\) governs the increase in scale of the  corresponding local shrinkage parameters. The larger the estimated values of \(\beta_{\text{epi}}\) are, the more likely it is to see non-zero regression coefficients for these HLA alleles.

\subsection*{Full model specification}

The full specification of the model is given by:

\begin{equation}
  \label{eq:full}
  \begin{aligned}
   y_{ik} &\sim \text{bernoulli}(\theta_{ik}) &\\
   \theta_{ik} &=
        \text{inv\_logit}(\beta_{0_{k}} + \gamma_{k}\text{logit}(P(y_{ik}=1|\Psi)) & \\
        \hspace{-6cm} &~~~~ +\sum_{k=1}^{D}X_{ij}\beta_{\text{HLA}_{jk}}) & \\
    \beta_{0_{k}} &\sim \text{Normal}(0, 100^2) & \text{(*)}\\
   \gamma_{k} &\sim \text{Normal}(\mu_{\text{phy}}, \sigma_{\text{phy}}^2) & \\
    \mu_{\text{phy}} &\sim \text{Normal}(1, 1) & \text{(*)} \\
   \sigma_{\text{phy}} &\sim \text{Normal}^{+}(0, 0.5) & \text{(*)} \\
   \beta_{\text{HLA}_{jk}} &\sim \text{Normal}(0, \tau_{k}^{2}\tilde{\lambda}^{2}_{jk}) & \\
   \tilde{\lambda}^{2}_{jk} &= \frac{c_{k}^{2}\lambda_{jk}^{2}}{c_{k}^{2} + \tau_{k}^{2}\lambda_{jk}^{2}} & \\
    c_{k}^2 &\sim \text{Inv-Gamma}(3.5, 3.5) & \text{(*)} \\
   \lambda_{jk} &\sim \text{Cauchy}^{+}(0, \sigma_{j}^2\text{exp}(Z_{jk}\beta_{\text{epi}})) & \\
   \tau_{k} &\sim \text{Cauchy}^{+}(0, \tau_{0k}) & \text{(*)} \\
   \tau_{0k} &= \frac{10}{D-10}\frac{2}{\sqrt{N}} & \\
  \end{aligned}
\end{equation}

The full model specification includes some aspects that were not covered in the previous sections. In particular, the parameters \(\gamma_{k}\) are modelled hierarchically, which allows information to be partially pooled across replacements. However, the model would also work reasonably well with just a single global parameter \(\gamma\). All other additions (\(\tilde{\lambda}_{jk}^{2}\) and \(\tau_{0k}\)) are explained in detail in \cite{Piironen2017}. Briefly, the addition of  \(\tilde{\lambda}_{jk}^{2}\) allows some regularization for the non-zero coefficients and the parameterization of \(\tau_{0k}\) allows to place a prior on the expected number of non-zero coefficients.

All models were implemented in Stan~\cite{Stan2021}. The Stan code is available online in two versions: One optimized for readability and one optimized for speed by utilizing Stan's multithreading and GPU capabilities.
A Julia~\cite{Bezanson2017} package is available to run the model on custom data. Due to restrictions of dependencies (MHCFlurry and RAxML-ng), HAMdetector is only available on Linux. We additionally provide Docker images to easily run the model on any operating system in a containerized environment, assuming Docker is installed.

\subsection*{Prior justification}

Prior distributions are labeled with an asterisk in equation~\ref{eq:full}. They are meant to be weakly informative, which means that the goal is to provide some modest amount of shrinkage while still allowing for large absolute values given enough support by the observed data. One exception to this are the intercepts \(\beta_{0_k}\), which are essentially flat because they are well identified by the data alone.

The hierarchical mean and standard deviation of the phylogeny coefficients \(\gamma_{k}\) are chosen to place most probability mass at values of \(\gamma_{k}\) of around 1. In the absence of any HLA effects, a parameter value of 1 would mean that the estimate for the probability of observing replacement \(k\) is identical to the probability based on the phylogenetic model. Intuitively speaking, We therefore treat the phylogeny as a baseline and any observations that cannot be attributed to phylogeny are attributed to the HLA alleles or noise.

The prior on \(c_{k}^{2}\) implies a Student-t prior with 7 degrees of freedom and a scale of 1 on the non-zero HLA regression coefficients \(\beta_{\text{HLA}_{jk}}\). A Student-t prior with these parameters is a reasonable default choice for logistic regression models. The theoretical justification behind this regularized variant of the horseshoe prior is explained in detail in~\cite{Piironen2017}.

The value of \(\tau_{0k}\) is chosen to imply 10 effective non-zero HLA regression coefficients per replacement. The rationale behind this parameterization is again outlined in~\cite{Piironen2017}. The value of 10 is approximately based on available HIV epitope maps. Please note that this is only an approximate statement about the general magnitude, i.e. we would expect maybe a couple of HLA alleles to be associated with a single replacement, not much more.
The model is also parameterized in a way that assumes an equal degree of sparsity across all alignment positions a priori. We also tried to model \(\tau_k\) hierarchically, but observed sampling issues due to the resulting unfavorable geometry of the posterior.

\section{Results} \label{sec:results}

To show that our model works well in different real-world scenarios, we run versions of our model on several datasets:

\begin{itemize}
  \item A large HIV dataset consisting of 1383 sequences, that were kindly provided by X and were also used in the Phylogenetic Dependency Network study~\cite{Carlson2008}. \hl{DaHa: Wer ist für diese Sequenzen der Ansprechpartner?}
  \item A set of 351 HIV sequences mostly spanning the Pol gene from the Arevir database~\cite{Roomp2006}.
  \item A set of 544 Hepatitis-B-Virus sequences that were kindly provided by Jörg Timm.
  \item A set of 104 Hepatitis-D-Virus sequences that were kindly provided by Michael Roggendorf.
  \item A set of 41 HIV sequences that were kindly provided by Rongge Yang.
\end{itemize}

For all sequences, we applied the following data preparation steps:

\begin{enumerate}
  \item For each dataset, the sequences were split into subsequences, either by protein or gene.
  \item If not already present in this format, sequences were translated into their amino acid representation.
  \item RAxML-NG~\cite{Kozlov2019} version 1.0.0 was used to generate a maximum likelihood phylogenetic tree for each gene/protein using the \texttt{--model GTR+G+I} option with all other parameters set to default values. When available, we preferred to use RNA or DNA sequences over amino acid sequences for this step.
\end{enumerate}

In order to evaluate each building block of HAMdetector, we applied 4 different Bayesian models to each dataset, each of which includes the previous model as a special case:
A standard logistic regression model, a logistic regression model with horseshoe prior, a model additionally including phylogeny and the full model with horseshoe prior, phylogeny and epitope prediction.
As a comparison to existing methods, we also applied Fisher's exact test and the Phylogenetic Dependency Network.

\subsection*{Convergence diagnostics}

\hl{DaHo: $\hat{R}$ und $n_{\mathrm{eff}}$ sollten in den Methods eingeführt werden}
For all the Bayesian models, Markov Chain Monte Carlo sampling was done using Stan~\cite{Stan2021}. All model fits showed no signs of inference issues. In total, 4 different chains with 1000 samples each were drawn from the posterior. The effective sample size for all model parameters exceeded 500, R-hat convergence diagnostic values were below 1.01 in all cases.

\subsection*{Posterior predictive checks}
\hl{DaHo: Sollte in den Methods eingeführt werden}
Bayesian modelling allows a simple yet effective strategy to test models: By simulating data under the model's assumptions and then comparing simulated and observed data, model misspecifications can be identified and highlight possible model improvements.

Here, we compare the calibration of the posterior predictive probability of observing the replacement to the actual observed data using calibration plots.
The plots are similar to the one showed in figure~\ref{fig:phylogeny-calibration} but this time do not evaluate the probabilities based on the phylogenetic model alone, but show the calibration of the posterior predictive probabilities based on the full model.

An example calibration plot of the HAMdetector model on the Core protein sequence data of the HBV dataset is shown in figure~\ref{fig:calibration}. Plots for the other models and datasets are shown in the supplementary.

\begin{figure}[!ht]
  \includegraphics[width=1\linewidth]{plots/phylogeny_calibration.pdf}
  \caption{Calibration plot for the HBV dataset. Calibration plots for the other datasets are shown in the supplementary.
  All observations are first sorted by increasing estimated probability \(P(y_{ik}=1)\) and then grouped into \(n\) bins.
  For each bin, the fraction of observations with \(y_{ik}=1\) (observed event percentage) is compared to the midpoint of each bin (the value in the center between the lowest and highest probability). The error bars show the cutpoints for each bin. If the probabilities are calibrated correctly, each dot is supposed to scatter closely around the diagonal line.}
  \label{fig:calibration}
\end{figure}

\subsection*{Comparison to a list of known epitopes}

When testing models, we prefer to evaluate them on real-world data whenever possible. 
However, comparing against real-world data is not straight forward when identifying HLA-associated mutations, because if a tool identifies an HLA-associated mutation which is not yet documented, this could be either because it is a false positive, or a yet unknown HLA-associated mutation that has not been documented before. 
Also, viewing every non-identified HLA-associated mutation that is listed in the literature as a false negative is not correct either, because documented escape mutations do not necessarily have to show in each sequence alignment.
A third issue when comparing models is that HLA-associated mutations are not necessarily binary, so the true positive / false negative framework does not work as well.

To circumvent these issues and still compare against real-world data, we chose the following strategy:
For each model, all evaluated replacements are ranked by decreasing confidence of being an HLA associated mutation. That is, for Bayesian models we calculate the probability of the corresponding regression coefficient exceeding zero \(P(\beta_{HLA} > 0)\), for Fisher's exact test we sort HLA allele - replacement pairs by increasing p-values and for the Phylogenetic Dependency Network, we rank the associations by increasing q-values.
Then, a list of known epitopes is used to create a plot of the cumulative number of replacements inside the boundary of a known epitope vs. the corresponding rank. The underlying assumption for this kind of model check is that we expect to see an enrichment of replacements that lie inside the region of known epitopes and that this enrichment is stronger for better performing models.

\begin{figure}[ht!]
  \includegraphics[width=1\linewidth]{plots/comparison.pdf}
  \caption{Number of associations inside the boundary of known epitopes vs. rank. For each model, the pairs of HLA alleles and replacements are sorted by decreasing confidence of being an HLA associated mutation. Then, each replacement is compared to a list of known epitopes and a line of the cumulative number of associations inside known epitopes vs. the current rank is plotted for each model. Better performing models are expected to have an enrichment of associations that are inside of known epitopes. 
  PhyloD: Phylogenetic Dependency Network, Model 1: simple logistic regression model with broad Student-t priors, Model 2: logistic regression model with horseshoe prior, Model 3: logistic regression model with horseshoe prior and phylogeny, Model 4: Full model with epitope prediction. The thin gray lines denote random permutations of the list of HLA allele - replacement pairs and act as a baseline.}
  \label{fig:comparison}
\end{figure}

Figure~\ref{fig:comparison} shows a comparison of different models on the HBe protein of the HBV dataset. Plots for all other datasets are shown in the supplementary, but show identical patterns.

All models show a strong enrichment of HLA-associated mutations compared to a randomly permutated list of HLA allele - replacement pairs (gray lines). Fisher's exact test performs about as well as the simple logistic regression model, as from a statistical perspective, they are both similar.
The horseshoe prior alone is a drastic improvement over Fisher's exact test and the logistic regression model with non-sparsifying priors, even though it does not include any external information. The benefit comes from including the assumption that most regression coefficients are expected to be very close to 0.
The logistic regression model with horseshoe prior works roughly as well as the Phylogenetic Dependency Network, which includes much more information.
Note that the line for the Phylogenetic Dependency Network is shorter because it only outputs associations with a q-value lower than 0.2 by default.

The remaining models are ranked in order of the number of additional features they include. Additionally including phylogeny is a small improvement over the model with the horseshoe prior, while adding epitope prediction provides another strong improvement.

Note that this model only uses epitope \textit{prediction} software and does not use any information of experimentally confirmed epitopes. These are here only used for model evaluation.

\subsection{Leave-one-out cross-validation}
\hl{DaHo: LOO auch in Methods einführen}
In addition to testing the model against experimentally validated epitopes, we also performed leave-one-out cross-validation. While the comparison against known epitopes compares inferences of the regression coefficients, leave-one-out cross-validation compares different models in regard to their predictive performance, i.e. how well they can predict the presence or absence of a replacement in a sequence based on the available data.

As the compared models are all special cases of the complete model, we can use cross-validation to gain insight on the relative importance of each aspect of the model.

Leave-one-out cross-validation has historically been challenging to apply for Bayesian models, because fitting a model once for each left out data-point is often infeasible because of computational constrains.
Recently, Pareto-smoothed importance sampling has gained popularity to perform approximate leave-one-out cross-validation for Bayesian models. This method  approximates the leave-one-out posterior by re-weighting existing draws from the posterior distribution and has the advantage of requiring to fit the model only once. Additionally, the method provides good diagnostics to identify cases where the approximation is unreliable~\cite{Vehtari2016}.

Several utility (or cost) functions are used in practice to compare the predictive performance of models. For frequentist approaches, classification accuracy or receiver operator characteristic curves (ROC curves) are most common.
For Bayesian models, expected log-predictive density (elpd) is used most frequently.

elpd is defined as \(\sum_{i=1}^{n}\text{log}(\int p(y_i|\theta)p(\theta|y_{i-1})d\theta))\), i.e. the average log predictive density of the observed data points based on the leave-one-out posterior distributions.
This has the advantage over other performance measures like classification accuracy, that it not only takes the location of the predictive distribution into account (i.e. the number of correct predictions), but also the width (i.e. how confident the model is in its predictions).
elpd cannot be computed for frequentist models because the parameters of the model are viewed as fixed values.

Table \ref{loo} shows a comparison of models with increasing complexity run on the HBe protein in the HDV dataset.

\begin{table}[h!]
  \renewcommand{\arraystretch}{1.3}
  \centering
  \caption{Differences in elpd for models with increasing model complexity (HBe protein of the HBV dataset).
  The numbers in column \(\text{elpd}_\text{diff}\) show increases in elpd compared to the previous model (higher is better), e.g. the difference in elpd of the model with horseshoe prior is 949.8 compared to the baseline model, and the difference in elpd for the model with phylogeny and horseshoe prior is 4440.9 compared to the model just with the horseshoe prior. All differences in elpd are larger than several times the estimated standard error (column \(\text{se}_\text{diff}\)), indicating that models which account for more information are expected to have a better predictive performance.}
  \vspace{0.5cm}
  \begin{tabular}{l|r|rll}
  \multicolumn{1}{l|}{} & \multicolumn{1}{c|}{\(\text{elpd}_\text{diff}\)} & \multicolumn{1}{c}{\(\text{se}_\text{diff}\)} &  &  \\ \cline{1-3}
  logistic regression (baseline) & 0.0  & 0.0 &  &  \\
  + horseshoe prior     & 949.8  & 65.2 &  &  \\
  + phylogeny & 4440.9 & 94.4 &  &  \\
  + epitope prediction & 63.1   & 18.9 &  & 
  \end{tabular}
  \label{loo}
\end{table}

Predictive accuracy increases as the models include more and more information.
Note that this is not due to overfitting, because leave-one-out cross-validation approximates the model performance on unseen data. Additional tables are shown in the supplementary, the results are consistent across all datasets.

Note that the model with horseshoe prior alone already has a much higher elpd than the standard logistic regression model, even though it does not use any additional external data. This is because including the sparsity assumption allows the model to better separate signal from noise and the uncertainty of the close-to-zero coefficients does not propagate into uncertainty of the estimated y.

Including phylogeny further improves model performance, as the assumption of independent and identically distributed data does not hold for sequence data that share a common phylogenetic history.

Inclusion of epitope prediction does not improve elpd as much as inclusion of phylogeny and the sparsity assumption. This is because while addition of sparsity and phylogeny has an effect on all replacements and samples, epitope prediction only influences those replacements which are restricted by a given HLA allele and only those samples which are annotated with the allele. However, inclusion of epitope prediction is highly useful for determining which HLA alleles are associated with the occurrence of a replacement, as shown in the previous section.

\subsection{A real-world example}

We had the opportunity to test our model against a set of associations identified by Fisher's exact test where some of them turned out to be false positives~\cite{Karimzadeh2019}. This data is rare because negative results are usually not published in the literature, but provide a useful way to access model performance.
A comparison with a preliminary version of HAMdetector was published recently. Here we provide a comparison with results from Fisher's exact test~\cite{Budeus2016} and the mature version of HAMdetector used in this work.

Fisher's exact test and  HAMdetector were applied to the same data (table~\ref{tab:false-positives}).
The posterior probabilities given by HAmdetector denote the probability that the corresponding regression coefficient is positive. Associations with strong support in our model have a posterior probability close to 1, associations with no support a probability close to 0.5 (corresponding to a regression coefficient centered around 0).

The results show low posterior probabilities for most of the false-positives and high posterior probabilities for most replacements that could be experimentally validated in IFN gamma production assays.
One important exception is P89T~-~HLA-B*37, which was experimentally validated but does not have strong support in our model.

\begin{table}[h!]
  \caption{List of HLA associated mutations that were identified using Fisher's exact test (FET). Experimental validation showed that 9 out of 15 replacements could not be experimentally validated (column “confirmed“). The data were re-analyzed using HAMdetector. Except for P89T, all validated replacements have high posterior probabilities, while the replacements that could not be experimentally validated have lower evidence of being HLA associated. Note that all replacement - HLA allele pairs had low p-values using Fisher's exact test.}
  \vspace{0.5cm}
  \begin{tabular}{c|c|c|c|c}
  \thead{replacement} & \thead{allele}  & \thead{p-value \\ (FET)} & \thead{post. prob. \\ (HAMdetector)} & \thead{confirmed} \\
  \hline
  S170N       & B*15 & \(3\cdot10^{-8}\)      & 0.99                                & +                        \\
  D101E       & B*37 & 0.0002                        & 0.96                                & +                        \\
  R105K       & B*27 & 0.0011                        & 0.93                                & +                        \\
  R139K       & B*41 & 0.0034                        & 0.91                                & +                        \\
  E47D        & B*18 & 0.0027                        & 0.89                                & +                        \\
  D33E        & B*13 & 0.0001                        & 0.86                                & -                        \\
  T134A       & A*68 & 0.0045                        & 0.82                                & -                        \\
  K43R        & B*13 & 0.0021                        & 0.77                                & -                        \\
  D47E        & A*30 & 0.0010                        & 0.76                                & -                        \\
  K113R       & B*13 & 0.0043                        & 0.76                                & -                        \\
  P89T        & B*37 & 0.0011                        & 0.75                                & +                        \\
  A107T       & B*14 & 0.0028                        & 0.70                                & -                        \\
  P49L        & A*30 & 0.0031                        & 0.63                                & -                        \\
  Q100L       & B*13 & 0.0018                        & 0.61                                & -                        \\
  D96E        & B*13 & 0.0035                        & 0.50                                & -                       
  \end{tabular}
  \label{tab:false-positives}
 \end{table}

\section{Discussion}

We present HAMdetector, a Bayesian model to identify HLA-associated mutations in a multiple sequence alignment. We propose to use models that include as much information as possible and have shown that including sparsity and epitope prediction software can achieve better performance than existing methods. The logistic regression backbone can be easily extended, which allows the model to also be used in other contexts, for example in identifying associations between sequence data and other features.

% \hl{DaHo: wir haben zwar mehrere Datensätze, gehen aber praktisch nicht auf die Resultate mit diesen Datensätzen ein. Wie besprochen sollte dieses Manuskript eher methodisch sein, d.h.~es ist ok, dass wir nicht *viel* auf die virologische/immunologische Seite eingehen. Aber, dass wir überhaupt nichts zu den Resultaten sagen, ist etwas zu wenig. Z.B. könnte man die Vielfalt der Datensätze nutzen um den Punkt zu machen, dass die Verbesserungen nicht von den Details des Virus abhängen. Vielleicht fällt Dir ja noch was dazu ein.}

\thispagestyle{empty} % Removes page numbering from the first page
%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------
\FloatBarrier

%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------
\phantomsection
\bibliographystyle{unsrt}
\bibliography{references}
%----------------------------------------------------------------------------------------

\end{document}